---
title: "Withdrawal in Harvard Courses"
author: "Taehwan Kim"
date: "4/6/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE)

# Load in libraries
library(fs)
library(readxl)
library(ggthemes)
library(janitor)
library(ggrepel)
library(scales)
library(tidyverse)

# Download enrollment data

download.file(url = "https://registrar.fas.harvard.edu/files/fas-registrar/files/class_enrollment_summary_by_term_3.22.19.xlsx",
              destfile = "enroll_19.xlsx",
              mode = "wb")

download.file(url = "https://registrar.fas.harvard.edu/files/fas-registrar/files/class_enrollment_summary_by_term_03.06.18.xlsx",
              destfile = "enroll_18.xlsx",
              mode = "wb")

# Read in enrollment statistics to environment

e_19 <- read_xlsx("enroll_19.xlsx", skip = 3) %>% 
  clean_names(case = "snake") %>%
  filter(!is.na(course_name))

e_18 <- read_xlsx("enroll_18.xlsx", skip = 3) %>% 
  clean_names(case = "snake") %>%
  filter(!is.na(course_name))

# Delete downloaded files

fs::file_delete(c("enroll_19.xlsx","enroll_18.xlsx"))
```

# Introduction

How has Harvard changed over small periods of time? In this project, I will be analyzing the changes in Harvard's course enrollment over a year time-frame.

# Withdrawing Courses

Every year, a few students withdraw from courses that they commit to because of a variety of reasons; that may just be students' personal issues. However, here, I show the classes whose withdrawal number has actually increased, noting that more people are dropping out as time has progressed.

```{r popular}

# Function to determine withdraw difference between years as increase, decrease, or no change

change_type <- function(n) {
  ifelse(n == 0, "No Change", ifelse(n > 0, "Increase", "Decrease"))
}

# Create dataframe containing both course enrollment data

enrollments <- inner_join(e_19, e_18, by = c("course_id"), suffix = c(".19", ".18"))

# Create dataframe with necessary data for graphing

withdraws <- enrollments %>% 
  
  # Choose only used variables
  
  select(course_id, course_title.19, u_grad.19, u_grad.18, withdraw.18, withdraw.19) %>% 
  
  #  Calculate changes in withdrawals and what kind of change
  
  mutate(withdraw_change = withdraw.19 - withdraw.18) %>% 
  mutate(withdrawals = change_type(withdraw_change)) %>% 
  
  # Look at only courses that are run in both years
  
  filter(u_grad.19 != 0) %>% 
  filter(u_grad.18 != 0) %>% 
  
  # Change order so plot order can be changed in scatterplot
  
  arrange(desc(withdrawals))

# Color codes of Decrease, Increase, No Change for graph
# Decrease is green because a decrease in withdraw is actually good

colors <- c("#139E34", "#B62222", "#C5C5C5")

# Graphing withdraw differences

withdraws %>% 
  ggplot(aes(x = u_grad.18, y = u_grad.19, size = abs(withdraw_change))) +
  
  # Create scatterplot
  
  geom_point(aes(color = withdrawals, alpha = 0.2)) +
  
  # Change scale of axes so graph is more dense
  
  scale_x_sqrt() +
  scale_y_sqrt() +
  
  # Label graph and axes
  
  labs(title = "Change in Number of Students Withdrawing",
       subtitle = "in Spring Harvard Courses in 2018 and 2019",
       caption = "Source: FAS Harvard Registrar",
       size = "Size of Change",
       color = "Change in Num. Withdraws") +
  xlab("Number of Undergraduates in 2018") +
  ylab("Number of Undergraduates in 2019") +
  
  # Change color for withdraw type 
  
  scale_color_manual(values = colors) +
  
  # Change graph to theme without grid lines because 
  # scale makes lines look disorganized

  theme_tufte() +

  # Override legend size and position
  
  theme(legend.position = "right", legend.key.size = unit(0.5, "cm")) +
  scale_alpha(guide = "none") +
  scale_size(guide = guide_legend(reverse = TRUE)) +
  
  # Annotate the courses whose withdraws have increased by at least 2
  
  geom_text_repel(data = subset(withdraws, abs(withdraw_change) >= 2), mapping = aes(label = course_title.19), size = 2)



```

